Thermodynamic fragments for magic states in quantum computation

### 2021-02-25 magic_paper.tex
More tidying

In particular, any $\sigma$--fragment, with $\sigma$ not full-rank, can be approximated by the $\sigma'$--fragment defined as a mixture of $\sigma$ with uniform noise, $\sigma' = (1-\epsilon)\sigma + \epsilon\frac{1}{d}\id$, for some infinitesimal $\epsilon > 0$, so that $\sigma'$ is full-rank and arbitrarily close to $\sigma$.
There exist many $\sigma$--fragments with $\sigma$ not full-rank, such as when $\sigma$ is a pure stabilizer, e.g. the zero state as depicted in~\cref{fig:zero}. 
Operations in such fragments include important stabilizer operations like the replacement channel defined as $\E(\rho) = \sigma$ for all states $\rho$.

@@

The area $\A_\sigma(\rho)$ between the curve $L_{\rho|\sigma}$ and the line $y=1$ is a resource monotone in the $\sigma$--fragment. 
This is clear because for any state conversion like~\cref{eq:sigma_conv}, the Lorenz curve $L_{\tau|\sigma}$ is lower than $L_{\rho|\sigma}$, hence $\A_\sigma(\E(\rho)) \leq \A_\sigma(\rho)$.
The exact form of the monotone is given in~\cref{app:areamono}
\nick{should we include the area? Have a look at~\cref{app:areamono}}

@@

@@

In the sections~\cref{sec:unital,sec:stab}, we study majorization constraints on magic distillation arising in different fragments. 
For this reason, we define the $n$-copy, $\epsilon$-noisy Strange state,
\begin{equation}
    \rho_{\rm{S}}(\epsilon)^{\otimes n} \coloneqq \left[ (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \frac{1}{3}\id \right]^{\otimes n},
\end{equation}
in $\O_\sigma$, where $\ket{\rm{S}}$ is defined in~\cref{fig:strange} and the induced noise is uniform. 
For $\epsilon=0$, this state is a qutrit of maximal sum-negativity / mana~\cite{cit:veitch2} and therefore acts as the ideal distiallation target, analogous to a Bell state in bipartite entangelement theory.

In~\cref{fig:lcs}, we illustrate the Lorenz curves of pure and noisy $n$-copy Strange states for $n$ up to $4$ in different $\sigma$--fragments.
As shown in~\cref{lem:lcmax}, the curve peaks equal $1 + \sn{\rho_{\rm{S}}(\epsilon)^{\otimes n}}$ independently of the $\sigma$--fragment.
\begin{figure}%
    \centering
    \subfigure[][]{%
    \label{fig:lcs_maxmix}%
    \includegraphics[height=3cm]{figs/lcs_maxmix.pdf}
    %\caption{Maximally mixed state $\frac{1}{3}\id$}%
    }\hspace{1pt}%
    \subfigure[][]{%
    \label{fig:lcs_thermal}%
    \includegraphics[height=3cm]{figs/lcs_thermal.pdf}
    %\caption{Zero state $\ketbra{0}{0}$}%
    }\\
    \subfigure[][]{%
    \label{fig:lcs_zero}%
    \includegraphics[height=3cm]{figs/lcs_zero.pdf}
    %\caption{Bound state}%
    }\hspace{1pt}%
    \subfigure[][]{%
    \label{fig:lcs_one}%
    \includegraphics[height=3cm]{figs/lcs_one.pdf}
    %\caption{Strange state $\ketbra{S}{S}$}%
    }
    \caption{\textbf{Lorenz curves of $n$-copy Strange states.} Lorenz curves of $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ for $n=1,2,3,4$.
    Solid lines represent pure Strange states ($\epsilon = 0$); dashed lines represent noisy Strange states ($\epsilon = 0.1$).
    \subref{fig:lcs_maxmix} Unital fragment; \subref{fig:lcs_thermal} Thermal fragment with $H = (0,1,2)$ and $\beta = 0.5$; \subref{fig:lcs_zero} Fragment $\O_{\ketbra{0}{0}}$; \subref{fig:lcs_one} Fragment $\O_{\ketbra{1}{1}}$.    
    \nick{change / move / remove ?}
    }%
    \label{fig:lcs}
\end{figure}

@@

In general, noisy circuits are well-described by the unital fragment.
To see this, consider incorporating noisy channels in the circuit, for example dephasing channels as in~\cref{eq:dephase} defined in different bases.
This process destroys the circuit symmetries, except for the invariance of the maximally mixed state.
Dephasing and bit-flip error channels are examples of the many error-inducing channels that respect the unital symmetry. \nick{expand}

@@

We consider the problem of purifying $n$ copies of a noisy Strange state $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ into a smaller number of copies $n'$ of a less noisy strange state $\rho_{\rm{S}}(\epsilon')^{\otimes n'}$, with $\epsilon' \leq \epsilon$ and $n' \leq n$,
\begin{equation}\label{eq:sudist}
	\rho_{\rm{S}}^{\otimes n} \longrightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n-n')},
\end{equation}
where all copies $n, n', n - n'$ are even.
Since the state $\frac{1}{3} \id$ is free, tensoring in copies of it does not affect the distillation process.

Since the initial state can be written in the form $\rho_{\rm{S}}(n, \epsilon, 0)$, it suffices to construct the Lorenz curve $\lc{\rho_{\rm{S}}}{\id/3}(x)$ for the generic noisy strange state $\rho_{\rm{S}}(n, \epsilon ,n')$.
This curve is defined at $9^n$ points between $0$ and $1$.
The exact expressions for the coordinates of these points can take $8$ different forms, depending on whether the noise level $\epsilon$ is greater than or less than $\frac{3}{7}$, the parity of the number of copies $n$ is even/odd and the location relative to the curve peak is on the left hand side (LHS -- including the maximum) / right hand side (RHS).
The full details are provided in~\cref{app:lcsu_technical}.

Here we focus on the case of Lorenz curves of even copies $n, n'$ and low noise levels ($\epsilon \leq 3/7$).
The part of the curve on the LHS of its peak is constructed by the positive components of the state Wigner distribution that take values $w_i$, with associated multiplicities $m_i$,
\begin{align}
	w_i &= \left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^{2i}\left( -\frac{1}{3} + \frac{4}{9}\epsilon \right)^{n-2i}, \\
	m_i &= 8^{2i}\binom{n}{2i},
\end{align}
where $i=0,\dots,\frac{n}{2}$.
From this, it is readily seen that $\lc{\rho}{\id/3}(x)$ has elbows on the LHS of its peak located at points $x=x_i$, where
\begin{equation}
	x_i = \Phi_+ \left( 2i; n , \frac{8}{9} \right),
\end{equation}
with $i = 0,1, \dots , n$ and where $\Phi_+(i,n,p)$ is an even-power cumulant distribution function given by
\begin{equation}
	\Phi_+(2i; n, p) \coloneqq \sum_{\ell=0}^{i} \binom{n}{2\ell} p ^{2\ell} (1-p)^{n-2\ell}.
\end{equation}
We provide technical details on the function $\Phi_+$ in the appendices.

@@

We now generalise the approach taken in the previous section and consider bounds on magic distillation for an arbitrary stabilizer fragment, $\R_\sigma$ where $\sigma$ is any quantum state $\sigma \in \rm{STAB}$. In other words we consider those bounds on distillation that apply when the free operations have $\sigma$ as a fixed point.

These bounds can be interpreted in two different ways: on one hand they can be viewed as a family of upper bounds parameterized by a stabilizer state $\sigma$, on the other we can associate $\sigma$ to actual hardware limitations or to biased noise models in which it is an equilibrium state of some kind. 
Without loss of generality we can always write $\sigma = \frac{1}{\Z} e^{-\beta H}$ for some $\beta \ge 0$ and Hamiltonian $H$ (either effective or actual)\footnote{Technicalities arise for the case where $\sigma$ is not full rank, but this can be still described via $ \beta \rightarrow 0$ limiting process.}.

Given this context, we can now state the following result, which is dependent on the free energy $F_\beta$ of a Gibbs state $\sigma$ where $ \beta F_\beta =  \beta\tr[H \sigma] - S(\sigma)$. We also fix the noisy Strange state to be of the form 
\begin{equation}
	\rho_{\rm{S}}(\epsilon) \coloneqq (1-\epsilon)\ketbra{\rm{S}} + \epsilon \frac{1}{3}\id,
\end{equation}
but more general states can be handled in exactly the same manner. In the following result we consider the bounds on distilling out pure states $|S\>\<S|$, with $\epsilon'=0$, but these bounds are easily generalised to purification bounds with $\epsilon' \ne 0$. Finally, we assume that $\epsilon \le 3/7$, so that the noisy magic state has \nick{a dominant negative component}.
\begin{theorem}Let $\sigma$ be a qutrit stabilizer state, given by $\sigma = \frac{1}{\Z} e^{-\beta H}$ with $H$ having eigenvalues $E_0 \le E_1 \le E_2$, and where $\beta$ is an (effective) inverse temperature for the state. 
We define
\begin{equation}
	\beta_\star \coloneqq \frac{1}{E_2 - E_0} \ln{2}
\end{equation}
and for $\beta \leq \beta_\star$ define
\begin{equation}
	\epsilon_{\star}(\beta) \coloneqq 3 - \dfrac{18}{8-e^{(E_2 - E_0)\beta}}.
\end{equation}
a threshold error rate. Then any distillation rate $R(\epsilon, \beta)$ in the $\sigma$--fragment of a magic theory is bounded as follows:

@@

@@

\subsubsection{Distillation bounds}

@@

There exist $d(d+1)$ pure stabilizer states for a $d$-dimensional system, all of which correspond to a uniform distribution over an affine line in the phase space \nick{CITE}.
As an example, the distribution of the qutrit zero state $\ketbra{0}$ is illustrated in~\cref{fig:zero}.
Any pure stabilizer state $\sigma$ is a rank--$1$ state and contains $d^2 - d$ zeros in its Wigner distribution, so we approximate it by the full--rank state $\sigma' = (1-\epsilon)\sigma + \epsilon \frac{1}{d}\id$ for some $\epsilon > 0$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\null\newpage
\section{Lower bounds in $\sigma$--fragments}
\ddd{[Am throwing down some rough material, to be polished later.]}

>>>>

In particular, any $\sigma$--fragment, with $\sigma$ not full-rank, can be approximated by the $\sigma'$--fragment defined as a mixture of $\sigma$ with uniform noise, $\sigma' = (1-\epsilon)\sigma + \epsilon\frac{1}{d}\id$, for some infinitesimal $\epsilon > 0$, so that $\sigma'$ is full-rank and arbitrarily close to $\sigma$.
Important examples include pure stabilzer states which are rank-$1$ states e.g. the zero state depicted in~\cref{fig:zero}. 
Operations in such fragments include important stabilizer operations like the replacement channel, $\E(\rho) = \sigma$ for all states $\rho$.

@@

The area $\A_\sigma(\rho)$ between the curve $L_{\rho|\sigma}$ and the line $y=1$ is a resource monotone in the $\sigma$--fragment. 
This is clear because for any state conversion like~\cref{eq:sigma_conv}, the Lorenz curve $L_{\tau|\sigma}$ is lower than $L_{\rho|\sigma}$, hence $\A_\sigma(\E(\rho)) \leq \A_\sigma(\rho)$.
The exact form of the monotone is given in~\cref{app:areamono}.
\nick{should we include the area? Have a look at~\cref{app:areamono}}

@@

\nick{Should this be a section on its own or a subsection of the previous section?}

@@

In the sections~\cref{sec:unital,sec:stab}, we study majorization constraints on magic distillation arising in different fragments. 
For this reason, we define the $n$-copy, $\epsilon$-noisy Strange state,
\begin{equation}\label{eq:noisysn}
    \rho_{\rm{S}}(\epsilon)^{\otimes n} \coloneqq \left[ (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \frac{1}{3}\id \right]^{\otimes n},
\end{equation}
in $\O_\sigma$, where $\ket{\rm{S}}$ is defined in~\cref{fig:strange} and the induced noise is uniform. 
For $\epsilon=0$, this state is a qutrit of maximal sum-negativity / mana~\cite{cit:veitch2} and therefore acts as the ideal distiallation target, analogous to a Bell state in bipartite entangelement theory.

In~\cref{fig:lcs}, we illustrate the Lorenz curves $\lc{\rho_{\rm{S}}}{\sigma}$\footnote{We drop parameters $\epsilon$ and $n$ for clarity} of pure and noisy $n$-copy Strange states for $n$ up to $4$ in different $\sigma$--fragments.
As shown in~\cref{lem:lcmax}, the curve peaks equal $1 + \sn{\rho_{\rm{S}}(\epsilon)^{\otimes n}}$ independently of the $\sigma$--fragment.
\begin{figure}%
    \centering
    \subfigure[][]{%
    \label{fig:lcs_maxmix}%
    \includegraphics[height=3cm]{figs/lcs_maxmix.pdf}
    %\caption{Maximally mixed state $\frac{1}{3}\id$}%
    }\hspace{1pt}%
    \subfigure[][]{%
    \label{fig:lcs_thermal}%
    \includegraphics[height=3cm]{figs/lcs_thermal.pdf}
    %\caption{Zero state $\ketbra{0}{0}$}%
    }\\
    \subfigure[][]{%
    \label{fig:lcs_zero}%
    \includegraphics[height=3cm]{figs/lcs_zero.pdf}
    %\caption{Bound state}%
    }\hspace{1pt}%
    \subfigure[][]{%
    \label{fig:lcs_one}%
    \includegraphics[height=3cm]{figs/lcs_one.pdf}
    %\caption{Strange state $\ketbra{S}{S}$}%
    }
    \caption{\textbf{Lorenz curves of $n$-copy Strange states.} Lorenz curves of $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ for $n=1,2,3,4$.
    Solid lines represent pure Strange states ($\epsilon = 0$); dashed lines represent noisy Strange states ($\epsilon = 0.1$).
    \subref{fig:lcs_maxmix} Unital fragment; \subref{fig:lcs_thermal} Thermal fragment with $H = (0,1,2)$ and $\beta = 0.5$; \subref{fig:lcs_zero} Fragment $\O_{\ketbra{0}{0}}$; \subref{fig:lcs_one} Fragment $\O_{\ketbra{1}{1}}$.    
    \nick{change / move / remove ?}
    }%
    \label{fig:lcs}
\end{figure}

@@

In general, noisy circuits are well-described by the unital fragment.
To see this, consider incorporating noisy channels in the circuit, for example dephasing channels as in~\cref{eq:dephase} defined in different bases.
This process destroys the circuit symmetries, except for the invariance of the maximally mixed state.
Dephasing and bit-flip error channels are examples of the many error-inducing channels that respect the unital symmetry. \nick{Expand on significance of unital fragment}

@@

We consider the problem of purifying $n$ copies of a noisy Strange state $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ as given in~\cref{eq:noisysn} into a smaller number of copies $n'$ of a less noisy strange state $\rho_{\rm{S}}(\epsilon')^{\otimes n'}$, with $\epsilon' < \epsilon$ and $n' \leq n$,
\begin{equation}\label{eq:sudist}
	\rho_{\rm{S}}(\epsilon)^{\otimes n} \longrightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n-n')},
\end{equation}
where all copies $n, n', n - n'$ are even.
Since the state $\id / 3$ is free, tensoring in copies of it does not affect the distillation process.

The Lorenz curve $\lc{\rho_{\rm{S}}}{\id/3}$, for some general noise parameter $\epsilon$ and number of copies $n$, is defined at $9^n$ points between $0$ and $1$.
The exact expressions for the coordinates of these points can take $8$ different forms, depending on whether the noise level $\epsilon$ is greater or less than $\frac{3}{7}$, the parity of the number of copies $n$ is even or odd and the location relative to the curve peak is on the left hand side (LHS -- including the curve peak) or right hand side (RHS) of the curve peak.
The full details for the Lorenz curve construction are provided in~\cref{app:lcsu_technical}.

\nick{In the following, should we talk about the coordinates of all elbows and then say that we are gonna use only the first elbow, or should we leave full-elbow analysis for the appendix and only talk about first elbow here?}

Here we focus on the case of Lorenz curves with even copies $n, n'$ and low noise levels ($\epsilon' < \epsilon \leq 3/7$).
The part of the curve on the LHS of its peak is constructed by the positive components of the state Wigner distribution that take values $w_i$, with associated multiplicities $m_i$,
\begin{align}
	w_i &= \left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^{2i}\left( -\frac{1}{3} + \frac{4}{9}\epsilon \right)^{n-2i}, \\
	m_i &= 8^{2i}\binom{n}{2i},
\end{align}
where $i=0,\dots,\frac{n}{2}$.
It is readily seen that $\lc{\rho_{\rm{S}}}{\id/3}(x)$ has elbows on the LHS of its peak located at points $x=x_i$, where
\begin{equation}
	x_i = \Phi_+ \left( 2i; n , \frac{8}{9} \right),
\end{equation}
with $i = 0,1, \dots , n$ and where $\Phi_+(i,n,p)$ is an even-power cumulant distribution function given by
\begin{equation}
	\Phi_+(2i; n, p) \coloneqq \sum_{\ell=0}^{i} \binom{n}{2\ell} p ^{2\ell} (1-p)^{n-2\ell}.
\end{equation}
We provide technical details on the function $\Phi_+$ in the appendices.

@@

We now generalise the approach taken in the previous section and consider bounds on magic distillation for an arbitrary stabilizer fragment, $\R_\sigma$ where $\sigma$ is any quantum state $\sigma \in \stab$. 
In other words, we consider those bounds on distillation that apply when the free operations have $\sigma$ as a fixed point.

These bounds can be interpreted in two different ways: on one hand they can be viewed as a family of upper bounds parameterized by a stabilizer state $\sigma$, on the other we can associate $\sigma$ to actual hardware limitations or to biased noise models in which it is an equilibrium state of some kind. 
Without loss of generality, we can always write a stabiler state $\sigma$ as a Gibbs state $\sigma = \frac{1}{\Z_\beta} e^{-\beta H}$ for some $\beta \geq 0$ and Hamiltonian $H$ as discussed in~\cref{sec:major_frag}.

We focus on the distillation process
\begin{equation}\label{eq:sudist}
	\rho_{\rm{S}}(\epsilon)^{\otimes n} \longrightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \gamma_\beta^{\otimes (n-n')},
\end{equation}
where the noisy Strange state is given in~\cref{eq:noisysn}, but more general states can be handled in exactly the same manner.
Notice that tensoring in copies of $\gamma_\beta$ does not affect the process, since the state is free.
In the following result, all copies $n, n', n - n'$ are even with $n > n'$, and $\epsilon'=0$, but the bounds are easily generalised to odd numbers of copies and $\epsilon'$ such that $0 < \epsilon' < \epsilon$. 
Finally, we assume a \nick{generous} range of initial noise levels, $\epsilon \leq 3/7$, so that the largest Wigner component of the 1-copy noisy magic state is negative.

\nick{Specify: How general can the states get? What about not full-rank states? Should we include the case of odd $n$ or non-zero $\epsilon'$ in the appendix? What about $\epsilon>0$?}

Given this context, we now provide the following result, which depends on the free energy $F_\beta$ of state $\sigma$,
\begin{equation}
	F_\beta \coloneqq \tr[H \sigma] - \beta^{-1}S(\sigma) = -\beta^{-1}\log{\Z_\beta},
\end{equation}
where the von Neumann entropy is $S(\sigma) \coloneqq -\tr[\sigma\log{\sigma}]$.

\begin{theorem}\label{thm:stab_bounds}
	Let $\sigma$ be a qutrit stabilizer state, given by $\sigma = \frac{1}{\Z} e^{-\beta H}$ with $H$ having eigenvalues $E_0 \le E_1 \le E_2$, and where $\beta$ is an (effective) inverse temperature for the state. 
We define
\begin{equation}
	\beta_\star \coloneqq \frac{1}{E_2 - E_0} \ln{2}
\end{equation}
and for $\beta \leq \beta_\star$ define
\begin{equation}
	\epsilon_{\star}(\beta) \coloneqq 3 - \dfrac{18}{8-e^{(E_2 - E_0)\beta}}.
\end{equation}
a threshold error rate. Then any distillation rate $R(\epsilon, \beta)$ in the $\sigma$--fragment of a magic theory is bounded as follows:

@@

\begin{proof}
	Here we provide a sketch of the proof with all the details included in \nick{the appendix}.

\nick{Rest of section needs tidying}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lower bounds in $\sigma$--fragments}
\ddd{[Am throwing down some rough material, to be polished later.]}
